### 2023-01-24 src/ms.tex
brought paper up to first compilable state

\documentclass[twocolumn,linenumbers]{src/aastex631}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}

\usepackage{amsmath}
\usepackage{cancel}
\usepackage{float}

\shorttitle{Self Consistently Modeling NGC 2808}
\shortauthors{Boudreaux et al.}
% \watermark{DRAFT}
\graphicspath{{./}{figures/}{src/figures}}

\begin{document}

\title{Chemically Self-Consistant Modeling of the Globular Cluster NGC 2808 and its Effects on the Inferred Helium abundance of Multiple Stellar Populations.}

\correspondingauthor{Thomas M. Boudreaux}
\email{thomas.m.boudreaux.gr@dartmouth.edu, thomas@boudreauxmail.com}

\author[0000-0002-2600-7513]{Thomas M. Boudreaux}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author[0000-0003-3096-4161]{Brian C. Chaboyer}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author[0000-0002-2012-7215]{Gregory Feiden}
\affiliation{Department of Physics and Astronomy, University of North Georgia, Dahlonega, GA 30533, USA}

\begin{abstract}
	Over its approximately 100 year history stellar modeling has become an
	essential tool for understanding certain astrophysical phenomena which are not
	directly observable. Modeling allows for empirical constraints --- such as
	elemental abundances, luminosities, and effective temperatures --- to strongly
	inform non-observables such as a star's age, mass, and radius. Here we propose a
	thesis in five parts, related through their use of both modeling and the
	Dartmouth Stellar Evolution Program (DSEP) to conduct this modeling. In two of
	the parts of this thesis we will use DSEP, in conjunction with atmospheric
	boundary conditions generated by collaborators, to build chemically
	self-consistent models of multiple populations (MPs) in the globular clusters
	NGC 2808, 47 Tuc, and NGC 6752. We will infer helium abundances across MPs and
	compare these inferred abundances to those from models which do not consider as
	careful a handling of a star's chemistry. PLACEHOLDER
\end{abstract}

\keywords{}

\input{src/sections/intro.tex}

\begin{acknowledgments}
	This work has made use of the NASA astrophysical data system (ADS). We
	would like to thank Elisabeth Newton and Aaron Dotter for their support and
	for useful disscusion related to the topic of this paper. We acknowledge
	the support of a NASA grant (No. 80NSSC18K0634). 
\end{acknowledgments}

% \software{
% 	The Dartmouth Stellar Evolution Program (DSEP) \citep{Dotter2008},
% 	\texttt{FreeEOS} \citep{Irwin2012},
% 	\texttt{pyTOPSScrape} \citep{Boudreaux22}
% }

\bibliography{src/bib/ms}{}
\bibliographystyle{aasjournal}

\end{document}