### 2023-01-26 src/sections/intro.tex
Started cleaning up introduction

\section{Introduction}\label{sec:Intro}
Globular clusters (GCs) are among the oldest observ- able objects in the
universe \citep{Pen11}. They are characterized by high densities with typical
half-light radii of $\le$10 pc \citep{Vanderburg2010}, and typical masses
ranging from $10^{4}$--$10^{5}$ M$_{\odot}$ \citep{Bro06} --- though some GCs
are signifigantly larger than these typical values {\color{red}EXAMPLE}. GCs
provide a unique way to probe stellar evolution \citep{Bau03}, galaxy formation
models \citep{Boy18,Kra05}, and dark matter halo struc- ture \citep{Hud18}.
{\color{red} BRING IN SOME MORE RECENT CITATIONS.}

Whereas, people have have often tried to categorized objects as GCs through
relations between half-light radius, density, and surface brightness profile,
in fact many objects which are generally thought of as GCs don't cleanly fit
into these cuts {\color{red} EXAMPLE + CITATION}. Consequently,
\citet{Carretta2010} proposed a definition of GC based on observed chemical
inhomogeneities in their stellar populations. The modern understanding of GCs
then is not simply one of a dense cluster of stars which may have chemical
inhomogeneities and multiple populations; rather, it is one where those
chemical inhomogeneities and multiple populations themselves are the defining
element of a GC.

All Milky Way globular clusters older than 2 Gyr studied in detail show
populations enriched in He, N, and Na while also being deplete in O and C
\citep{Piotto2015,Bastian2018}. These light element abundance patterns also are
not strongly correlated with variations in heavy element abundance, resulting
in spectroscopically uniform Fe abundances between populations. Further,
high-resolution spectral studies reveal anti-correlations between N-C
abundances, Na-O abundances, and potentially Al-Mg \citep{Sneden1992,
Gratton2012}. Typical stellar fusion reactions can deplete core oxygen;
however, the observed abundances of Na, Al, and Mg cannot be explained by the
likes of the CNO cycle \citep{Prantzos2007}.

Formation channels for these multiple populations remain a point of debate
among astronomers. Most proposed formation channels consist of some older,
more massive, population of stars polluting the pristine cluster media before a
second population forms, now enriched in heavier elements which they themselves could
not have generated \citep[for a detailed review see ][]{Gratton2012}. The four
primary candidates for these polluters are asymptotic giant branch stars
\citep[AGBs,][]{Ventura2001,DErcole2010}, fast rotating massive stars
\citep[FRMSs,][]{Decressin2007}, super massive stars
\citep[SMSs,][]{Denissenkov2014}, and massive interacing binaries
\citep[MIBs,][]{deMink2009, Bastian2018}.

Hot hydrogen burning (proton capture), material transport to the surface, and
material ejection into the intra-cluster media are features of each of these
models and consequently they can all be made to {\it qualitatively} agree with
the observed elemental abundances. However, none of the standard models can
currently account for all specific abundances \citep{Gratton2012}. AGB and FRMS
models are the most promising; however, both models have difficulty reproducing
severe O depletion \citep{Ventura2009,Decressin2007}. Moreover, AGB and FRMS
models require signifigant mass loss ($\sim 90\%$) between cluster formation
and the current epoch --- implying that a signifigant fraction of halo stars
formed in GCs \citep{Renzini2008,DErcole2008,Bastian2015}.

In addition to the light-element anti-correlations observed it is also known
that younger populations are signifigantly enhanced in Helium
\citep{Piotto2007, Piotto2015, Latour2019}. Depending on the cluster, Helium
mass fractions as high as $Y=0.4$ have been inferred \citep[e.g][]{Milone2015}.
However, due to the relatively high and tight temperature range of partial
ionization for He it cannot be observed in globular clusters; consequently, the
evidence for enhanced He in GCs originates from comparison of theoretical
stellar isochrones to the observed color-magnitude-diagrams of globular
clusters. Therefore, a careful handling of chemistry is essential when modeling
with the aim of discriminating between MPs; yet, only a very limited number of
GCs have yet been studied with chemically self-consistent (structure and
atmosphere) isochrones \citep[e.g.][NGC 6752]{Dotter2015}.

This thesis will contain chapters where we expand the number of clusters which
have been self-consistently modeled. In this chapter we will focus on
chemically self-consistent modeling of the two extreme population of NGC 2808
identified by \citep{Milone2015}, A and E.

One key element of NGC 2808 modeling is the incorporation of new atmospheric
models, generated from the \texttt{MARCS} grid of model atmospheres \citep{Plez2008},
which match interior elemental abundances. \texttt{MARCS} provides one-dimensional,
hydrostatic, plane-parallel and spherical LTE atmopsheric models
\citep{Gustafsson2008}. Members of our collaboration have generated atmospheric 
models for populations A and E.  Integration of these new model atmospheres
into DSEP is ongoing.

\subsubsection{Population Opacities}
For similar reasons as discussed in \S\ref{sec:p1} we conduct this research
with OPLIB high-temperature opacity tables as opposed to OPAL tables. We will
also generate low temperature opacity tables using the \texttt{MARCS}.
Moreover, we confirm that the atmosphere and structure meet in an optically
thick region of the star by shifting the atmospheric fitting point from an optical
depth of $\tau = 2/3$ (used by DSEP currently for \texttt{PHOENIX} model atmospheres) to
some higher $\tau$. We will experiment to identify the best optical depth to
fit at..

These population have been studied in depth by Feiden and their chemical
compositions were determined in \citet{Milone2015} (see Table 2 in that paper).
While we cannot yet evolve DSEP models with these new boundary conditions, we
can make a first pass investigation of the affect of OPLIB opacities (Figure
\ref{fig:NGC2808ISO}). Note how the models generated using OPLIB opacity tables
have a systematically lower luminosity. This discrepancy is consistent with
the overall lower opacities of the OPLIB tables.

\subsubsection{Additional Consistency}
The isochrones generally used to infer the degree of helium enhancements assume that
convection operates in the same manner in metal-poor stars as it does in the
Sun. However, observations from \textit{Kepler} of metal-poor red giants
\citep{Bonaca2012, tayar2017correlation}, in concert with interferometric
radius determination of the metal-poor sub-giant HD 140283
\citep{creevey2015benchmark}, have shown that the efficiency of convection
changes with iron content. As the final portion of our work to more carefully
handle a star's chemistry, we will modify DSEP to capture this variation in
convective efficiency.