### 2023-06-30 src/ms.tex
Paper updates

\documentclass[twocolumn,linenumbers]{src/aastex631}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand\fidanka{\texttt{Fidanka} }

\usepackage{amsmath}
\usepackage{cancel}
\usepackage{float}

\shorttitle{Self Consistently Modeling NGC 2808}
\shortauthors{Boudreaux et al.}
% \watermark{DRAFT}
\graphicspath{{./}{figures/}{src/figures}}

\begin{document}

\title{Chemically Self-Consistant Modeling of the Globular Cluster NGC 2808 and its Effects on the Inferred Helium abundance of Multiple Stellar Populations.}

\correspondingauthor{Emily M. Boudreaux}
\email{emily.m.boudreaux.gr@dartmouth.edu, emily@boudreauxmail.com}

\author[0000-0002-2600-7513]{Emily M. Boudreaux}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author[0000-0003-3096-4161]{Brian C. Chaboyer}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author{Rehnata Hoh}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author[0000-0002-2012-7215]{Gregory Feiden}
\affiliation{Department of Physics and Astronomy, University of North Georgia, Dahlonega, GA 30533, USA}

\begin{abstract}
	Globular Clusters (GCs) provide a unique astrophysical laboratory for
	studying the formation and evolution of stars. GCs are old, dense, and it
	has historically been believed that they have a single stellar population
	(SSP, in some publications also simple stellar population). However, in the
	last two decades, it has been definitively shown that most if not all Milky
	Way GCs have multiple stellar populations (MPs). These MPs are chemically
	distinct from one another, primarily separated by light element abundance
	variations without the standard accompanying heavy element abundance
	variations. In addition to the light element abundance variations, MPs also
	have different inferred helium abundances. The origin of these MPs remains
	an open question, and one which is sensitive to the population - population
	compositional differences; consequently, the extent of the composition
	variations is a key parameter to constrain. Metal abundances may be
	directly measured spectroscopically; however, helium abundances are not
	directly observable in GCs. Instead, helium abundances are inferred from
	stellar models. It is therefore important to understand build stellar
	models that are self-consistent in the compositions of the structure,
	atmosphere, and opacity. In this work we present the first chemically
	self-consistent stellar models of the Milky Way Globular Cluster NGC 2808.
	We find that the helium abundance of the second generation of stars is
	higher than the first generation by {\color{red} SOME AMOUNT}.
	{\color{blue} This is in agreement with previous studies of NGC 2808.}
\end{abstract}

\keywords{Globular Clusters (656), Stellar evolutionary models (2046)}

\input{src/sections/intro.tex}

\input{src/sections/AtmopshericModels.tex}

\input{src/sections/modeling.tex}

\input{src/sections/fidanka.tex}

\input{src/sections/fitting.tex}

\input{src/sections/results.tex}

% \input{src/sections/discussion.tex}

\input{src/sections/conclusion.tex}

\begin{acknowledgments}
	This work has made use of the NASA astrophysical data system (ADS). We
	would like to thank Elisabeth Newton and Aaron Dotter for their support and
	for useful disscusion related to the topic of this paper. Additionally, we
	would like to thank Kara Fagerstrom, Aylin Garcia Soto, and Keighley
	Rockcliffe for their useful disscusion related to in this work. We
	acknowledge the support of a NASA grant (No. 80NSSC18K0634). 
\end{acknowledgments}

% \software{
% 	The Dartmouth Stellar Evolution Program (DSEP) \citep{Dotter2008},
% 	\texttt{FreeEOS} \citep{Irwin2012},
% 	\texttt{pyTOPSScrape} \citep{Boudreaux22}
% }

\bibliography{src/bib/ms}{}
\bibliographystyle{aasjournal}

\end{document}