### 2023-09-01 src/sections/modeling.tex

\section{Stellar Models}\label{sec:modeling}
We use the Dartmouth Stellar Evolution Program \citep[DSEP, ][]{Dotter2008} to
generate stellar models. DSEP is a well-tested, one-dimensional stellar
evolution code which includes a mixing length model of convection,
gravitational settling, and diffusion.

We use DSEP to evolve stellar models ranging in mass from 0.3 to 2.0 solar
masses from the zero-age main sequence (ZAMS) to the tip of the red giant
branch. Below 0.7 $M_{\odot}$ we evolve a model every 0.03 $M_{\odot}$ and
above 0.7 $M_{\odot}$ we evolve a model every 0.5 $M_{\odot}$. Additionally, we
evolve models over a grid of mixing length parameters, $\alpha_{MLT}$, from
$\alpha_{MLT} = 1.0$ to $\alpha_{MLT} = 2.0$ in steps of 0.1. In addition to the
mixing length grid the evolved grid of models also has dimensions population (A
or E) (Table \ref{tab:comp}) and helium abundance ($Y=0.24, 0.27, 0.3, 0.33, 0.36, 0.39$). Each model is evolved in DSEP's ``high
resolution'' mode and had a maximum allowed time step of 50 Myr.

\begin{deluxetable}{c|cc||c|cc}\label{tab:comp}

%% Keep a portrait orientation

%% Over-ride the default font size
%% Use Default (12pt)

%% Use \tablewidth{?pt} to over-ride the default table width.
%% If you are unhappy with the default look at the end of the
%% *.log file to see what the default was set at before adjusting
%% this value.

%% This is the title of the table.
\tablecaption{Population Composition}

%% This command over-rides LaTeX's natural table count
%% and replaces it with this number.  LaTeX will increment 
%% all other tables after this table based on this number
\tablenum{1}

%% The \tablehead gives provides the column headers.  It
%% is currently set up so that the column labels are on the
%% top line and the units surrounded by ()s are in the 
%% bottom line.  You may add more header information by writing
%% another line between these lines. For each column that requries
%% extra information be sure to include a \colhead{text} command
%% and remember to end any extra lines with \\ and include the 
%% correct number of &s.
  \tablehead{\colhead{Element} & \colhead{Pop A} & \colhead{Pop E} & \colhead{Element} & \colhead{Pop A} & \colhead{Pop E} 
}

%% All data must appear between the \startdata and \enddata commands
\startdata
Li & -0.08 & --- & In & -1.46 & --- \\
Be & 0.25 & --- & Sn & -0.22 & --- \\
B & 1.57 & --- & Sb & -1.25 & --- \\
C & 6.87 & 5.91 & Te & -0.08 & --- \\
N & 6.42 & 6.69 & I & -0.71 & --- \\
O & 7.87 & 6.91 & Xe & -0.02 & --- \\
F & 3.43 & --- & Cs & -1.18 & --- \\
Ne & 7.12 & 6.7 & Ba & 1.05 & --- \\
Na & 5.11 & 5.7 & La & -0.03 & --- \\
Mg & 6.86 & 6.42 & Ce & 0.45 & --- \\
Al & 5.21 & 6.61 & Pr & -1.54 & --- \\
Si & 6.65 & 6.77 & Nd & 0.29 & --- \\
P & 4.28 & --- & Pm & -99.0 & --- \\
S & 6.31 & 5.89 & Sm & -1.3 & --- \\
Cl & -1.13 & 4.37 & Eu & -0.61 & --- \\
Ar & 5.59 & 5.17 & Gd & -1.19 & --- \\
K & 3.9 & --- & Tb & -1.96 & --- \\
Ca & 5.21 & --- & Dy & -1.16 & --- \\
Sc & 2.02 & --- & Ho & -1.78 & --- \\
Ti & 3.82 & --- & Er & -1.34 & --- \\
V & 2.8 & --- & Tm & -2.16 & --- \\
Cr & 4.51 & --- & Yb & -1.42 & --- \\
Mn & 4.3 & --- & Lu & -2.16 & --- \\
Fe & 6.37 & --- & Hf & -1.41 & --- \\
Co & 3.86 & --- & Ta & -2.38 & --- \\
Ni & 5.09 & --- & W & -1.41 & --- \\
Cu & 3.06 & --- & Re & -2.0 & --- \\
Zn & 2.3 & --- & Os & -0.86 & --- \\
Ga & 0.78 & --- & Ir & -0.88 & --- \\
Ge & 1.39 & --- & Pt & -0.64 & --- \\
As & 0.04 & --- & Au & -1.34 & --- \\
Se & 1.08 & --- & Hg & -1.09 & --- \\
Br & 0.28 & --- & Tl & -1.36 & --- \\
Kr & 0.99 & --- & Pb & -0.51 & --- \\
Rb & 0.26 & --- & Bi & -1.61 & --- \\
Sr & 0.61 & --- & Po & -99.0 & --- \\
Y & 1.08 & --- & At & -99.0 & --- \\
Zr & 1.45 & --- & Rn & -99.0 & --- \\
Nb & -0.8 & --- & Fr & -99.0 & --- \\
Mo & -0.38 & --- & Ra & -99.0 & --- \\
Tc & -99.0 & --- & Ac & -99.0 & --- \\
Ru & -0.51 & --- & Th & -2.2 & --- \\
Rh & -1.35 & --- & Pa & -99.0 & --- \\
Pd & -0.69 & --- & U & -2.8 & --- \\
\enddata

%% Include any \tablenotetext{key}{text}, \tablerefs{ref list},
%% or \tablecomments{text} between the \enddata and 
%% \end{deluxetable} commands

%% General table comment marker
\tablecomments{Relative Metal composition used where a(H) = 12. Where the relative composition is the the same for both populations A and E it is only listed in the population A colum for the sake of visual clarity.}

%% General table references marker
\tablerefs{\citet{Milone2015}}

\end{deluxetable}

For each combination of population, $Y$, and $\alpha_{MLT}$ we use the
isochrone generation code first presented in \citet{Dotter2016} to generate a
grid of isochrones. The isochrone generation code identified equivalent
evolutionary points (EEPs) over a series of masses and interpolates between
them. {\color{blue} The grid of isochrones generated for this work is avalible
as a digital supplement to this paper.}