### 2023-12-10 src/ms.tex
Incorperated Brian's and Greg's notes

\documentclass[twocolumn,linenumbers]{src/aastex631}
\newcommand{\vdag}{(v)^\dagger}
\newcommand\aastex{AAS\TeX}
\newcommand\latex{La\TeX}
\newcommand\fidanka{\texttt{Fidanka} }

\usepackage{amsmath}
\usepackage{cancel}
\usepackage{float}

\shorttitle{Self Consistently Modeling NGC 2808}
\shortauthors{Boudreaux et al.}
% \watermark{DRAFT}
\graphicspath{{./}{figures/}{src/figures}}

\begin{document}

\title{Chemically Self-Consistant Modeling of the Globular Cluster NGC 2808 and its Effects on the Inferred Helium abundance of Multiple Stellar Populations.}

\correspondingauthor{Emily M. Boudreaux}
\email{emily.m.boudreaux.gr@dartmouth.edu, emily@boudreauxmail.com}

\author[0000-0002-2600-7513]{Emily M. Boudreaux}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author[0000-0003-3096-4161]{Brian C. Chaboyer}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author{Amanda Ash}
\affiliation{Department of Physics and Astronomy, University of North Georgia, Dahlonega, GA 30533, USA}

\author{Rehnata Hoh}
\affiliation{Department of Physics and Astronomy, Dartmouth College, Hanover, NH 03755, USA}

\author[0000-0002-2012-7215]{Gregory Feiden}
\affiliation{Department of Physics and Astronomy, University of North Georgia, Dahlonega, GA 30533, USA}

\begin{abstract}
  The Helium abundances in the multiple populations which are now known to
  comprise all closely studied Milky Way globular clusters are often inferred
  by fitting isochohrones generated from stellar evolutionary models to
  globular cluster photometry. It is therefore important to build stellar
  models that are chemically self-consistent in terms of their structure,
  atmosphere, and opacity. In this work we present the first chemically
  self-consistent stellar models of the Milky Way Globular Cluster NGC 2808
  using MARCS model atmospheres, OPLIB high-temperature radiative opacities,
  and AESOPUS low-temperature radiative opacities. These stellar models were
  fit to the NGC 2808 photometry using \fidanka, a new software tool that was
  developed optimally fit cluster photometry to isochrones and for population
  synthesis.  \fidanka can determine, in a relatively unbiased way, the ideal
  number of distinct populations which exist within a dataset and then fits
  isochrones to each population. We achieve this through a combination of
  Bayesian Gaussian Mixture Modeling and a novel number density estimation
  algorithm. Using \fidanka and F275W-F814W photometry from the Hubble UV
  Globular Cluster Survey we find that the helium abundance of the second
  generation of stars in NGC 2808 is higher than the first generation by
  $15\pm3\%$. This is in agreement with previous studies of NGC 2808.
\end{abstract}
% \begin{abstract}
% 	Globular Clusters (GCs) provide a unique astrophysical laboratory for
% 	studying the formation and evolution of stars. GCs are old, dense, and it
% 	has historically been believed that they have a single stellar population. However, in the
% 	last two decades, it has been definitively shown that most if not all Milky
% 	Way GCs have multiple stellar populations (MPs). These MPs are chemically
% 	distinct from one another, primarily separated by light element, including helium, abundance
% 	variations without the standard accompanying heavy element abundance
% 	variations. As the precise formation channel of these MPs remains
% 	an open question, and one which is sensitive to the population - population
% 	compositional differences, the extent of the composition
% 	variations between MPs is a key parameter to constrain. Many metal abundances may be
% 	directly measured spectroscopically; however, helium abundances are not
% 	directly observable in GCs. Instead, helium abundances are inferred from
% 	stellar models. It is therefore important to understand build stellar
% 	models that are self-consistent in the compositions of the structure,
% 	atmosphere, and opacity. In this work we present the first chemically
% 	self-consistent stellar models of the Milky Way Globular Cluster NGC 2808 using MARCS model atmospheres, OPLIB high-temperature radiative opacities, and AESOPUS low-temperature radiative opacities.
% 	We find that the helium abundance of the second generation of stars is
% 	higher than the first generation by {\color{red} SOME AMOUNT}.
% 	{\color{blue} This is in agreement with previous studies of NGC 2808.}
% \end{abstract}

\keywords{Globular Clusters (656), Stellar evolutionary models (2046)}

\input{src/sections/intro.tex}

\input{src/sections/AtmopshericModels.tex}

\input{src/sections/modeling.tex}

\input{src/sections/fidanka.tex}

\input{src/sections/fitting.tex}

% \input{src/sections/results.tex}

% \input{src/sections/discussion.tex}

\input{src/sections/conclusion.tex}

\begin{acknowledgments}
	This work has made use of the NASA astrophysical data system (ADS). We
	would like to thank Elisabeth Newton and Aaron Dotter for their support and
	for useful disscusion related to the topic of this paper. Additionally, we
	would like to thank Kara Fagerstrom, Aylin Garcia Soto, and Keighley
	Rockcliffe for their useful disscusion related to in this work. We
	acknowledge the support of a NASA grant (No. 80NSSC18K0634). 
\end{acknowledgments}

% \software{
% 	The Dartmouth Stellar Evolution Program (DSEP) \citep{Dotter2008},
% 	\texttt{FreeEOS} \citep{Irwin2012},
% 	\texttt{pyTOPSScrape} \citep{Boudreaux22}
% }

\bibliography{src/bib/ms}{}
\bibliographystyle{aasjournal}

\end{document}