### 2024-02-03 ArXiv/intro.tex

\section{Introduction}\label{sec:Intro}
Globular clusters (GCs) are among the oldest observable objects in the universe
\citep{Pen11}. They are characterized by high densities with typical half-light
radii of $\le$10 pc \citep{Vanderburg2010}, and typical masses ranging from
$10^{4}$--$10^{5}$ M$_{\odot}$ \citep{Bro06} --- though some GCs are
significantly larger than these typical values \citep[e.g. $\omega$ Cen,
][]{Richer1991}. GCs provide a unique way to probe stellar evolution
\citep{Bau03}, galaxy formation models \citep{Boy18,Kra05}, and dark matter
halo structure \citep{Hud18}.

The traditional view of Globular Clusters was that they consisted of a single
stellar population (SSP, in some publications this is referred to as a Simple
Stellar Population). This view was supported by spectroscopically uniform heavy
element abundances \citep{Carretta2010, Bastian2018} across most clusters (M54
and $\omega$Cen are notable exceptions, see \citet{Marino2015} for further
details), and the lack of evidence for multiple stellar populations (MPs) in
past color-magnitude diagrams of GCs \citep[i.e.][]{Sandage1953, Alcaino1975}.
However, over the last 40 years non-trivial star-to-star light-element
abundance variations have been observed \citep[i.e.][]{Smith1987} and, in the
last two decades, it has been definitively shown that most if not all Milky Way
GCs have MPs \citep{Gratton2004, Gratton2012, Piotto2015}. The lack of
photometric evidence for MPs prior to the 2000, can be attributed to the more
narrow color bands available, until very recently, to ground based photometric
surveys \citep{Milone2017}.

The prevalence of multiple populations in GCs is so distinct that the proposed
definitions for what constitutes a globular cluster now often center the
existence of MPs \citep[e.g.][]{Carretta2010}. Whereas, people have have often tried to categorized objects
as GCs through relations between half-light radius, density, and surface
brightness profile, in fact many objects which are generally thought of as GCs
don't cleanly fit into these cuts \citep{Peebles1968, Brown1991, Brown1995, Bekki2002}.
Consequently, \citet{Carretta2010} proposed a definition of GC based on
observed chemical inhomogeneities in their stellar populations. The modern
understanding of GCs then is not simply one of a dense cluster of stars that
may have chemical inhomogeneities and multiple populations; rather, it is one
where those chemical inhomogeneities and multiple populations themselves are
the defining element of a GC.

All Milky Way globular clusters older than 2 Gyr studied in detail show
populations enriched in He, N, and Na while also being deplete in O and C
\citep{Piotto2015,Bastian2018}. These light element abundance patterns also are
not strongly correlated with variations in heavy element abundance, resulting
in spectroscopically uniform Fe abundances between populations. Further,
high-resolution spectral studies reveal anti-correlations between N-C
abundances, Na-O abundances, and potentially Al-Mg \citep{Sneden1992,
Gratton2012}. Typical stellar fusion reactions can deplete core oxygen;
however, the observed abundances of Na, Al, and Mg cannot be explained by the
CNO cycle \citep{Prantzos2007}. Consequently, globular cluster populations must
be formed by some novel means.

Formation channels for these multiple populations remain a point of debate
among astronomers. Most proposed formation channels consist of some older,
more massive, population of stars polluting the pristine cluster media before a
second population forms, now enriched in heavier elements which they themselves could
not have generated \citep[for a detailed review see ][]{Gratton2012}. The four
primary candidates for these polluters are asymptotic giant branch stars
\citep[AGBs,][]{Ventura2001,DErcole2010}, fast rotating massive stars
\citep[FRMSs,][]{Decressin2007}, super massive stars
\citep[SMSs,][]{Denissenkov2014}, and massive interacting binaries
\citep[MIBs,][]{deMink2009, Bastian2018}.

Hot hydrogen burning (i.e. proton capture), material transport to the surface, and
material ejection into the intra-cluster media are features of each of these
models and consequently they can all be made to {\it qualitatively} agree with
the observed elemental abundances. However, none of the standard models can
currently account for all specific abundances \citep{Gratton2012}. AGB and FRMS
models are the most promising; however, both models have difficulty reproducing
severe O depletion \citep{Ventura2009,Decressin2007}. Moreover, AGB and FRMS
models require significant mass loss ($\sim 90\%$) between cluster formation
and the current epoch --- implying that a significant fraction of halo stars
formed in GCs \citep{Renzini2008,DErcole2008,Bastian2015}.

In addition to the light-element anti-correlations observed, it is also known
that younger populations are significantly enhanced in Helium
\citep{Piotto2007, Piotto2015, Latour2019}. Depending on the cluster, helium
mass fractions as high as $Y=0.4$ have been inferred \citep[e.g][]{Milone2015}.
However, due to both the relatively high and tight temperature range of partial
ionization for He and the efficiency of gravitational settling in core helium
burning stars, the initial He abundance of globular cluster stars cannot be
observed; consequently, the evidence for enhanced He in GCs originates from
comparison of theoretical stellar isochrones to the observed
color-magnitude-diagrams of globular clusters. Therefore, a careful handling of
chemistry is essential when modeling with the aim of discriminating between
MPs; yet, only a very limited number of GCs have been studied with
chemically self-consistent (structure and atmosphere) isochrones
\citep[e.g.][NGC 6752]{Dotter2015}.

NGC 2808 is the prototype globular cluster to host Multiple Populations.
Various studies since 2007 have identified that it may host anywhere from 2-5
stellar populations. These populations have been identified both
spectroscopically \citep[i.e.][]{Carretta2004, Carretta2006, Carretta2010, Gratton2011, Carretta2015, Hong2021} and photometrically \citep[i.e.][]{Piotto2007, Piotto2015, Milone2015, Milone2017, Pasquato2019}. Note
that recent work \citep{Valle2022} calls into question the statistical
significance of the detections of more than 2 populations in the spectroscopic
data. Here we present new, chemically self-consistent modeling of the
photometry of the two extreme populations of NGC 2808 identified by
\citet{Milone2015}, populations A and E. We use archival photometry from the
Hubble UV Globular Cluster Survey (HUGS) \citep{Piotto2015, Milone2017} in the
F275W and F814W passbands to characterize multiple populations in NGC 2808
\citep{Milone2015, Milone2015b}. Additionally, we present a likelihood analysis of the
photometric data of NGC 2808 to determine the number of populations present in
the cluster.