### 2024-03-22 refResponse/response.txt
Started work on ref report

# Response to the referee for the paper titled
# Chemically Self-Consistant Modeling of the Globular Cluster NGC 2808 and its Effects on the Inferred Helium abundance of Multiple Stellar Population.

We thank the referee for their careful reading and critique of our work. Below
we have attached a response to the referees comments enumerated in the same
manner in which we recived the report. All changes to the text of the manuscript
have been placed in boldface.

# Comment 1 [Chromosome Maps]
Referee:
  Milone et al. 2015 and other groups identified five distinct populations in NGC
  2808. To do so they used the chromosome map, a color-color combination of
  filters (including F814W and F275W, the two filters used in this study). This
  diagnostic tool was introduced to maximize the separation between stars with
  different light-element abundances (different filters are in fact sensitive to
  different light-elements), information which, as demonstrated, can only be
  partially retrieved from a (F275W-F814W, F814W) CMD. The authors do not justify
  the choice to use the (F275W-F814W, F814W) CMD, instead of the chromosome map.
  Furthermore, although the authors only find two populations from their
  analysis, they do not comment on why this is the case and, importantly, how
  this compares with what Milone et al. 2015 and others found. Finally, it is
  neither clear nor physically motivated why the method used in their work should
  be a better method for identifying multiple populations than the chromosome
  map. This is a crucial point and is completely missing.
Response:
  Overall, in response to this point, we have added text to the paper to
  summarize why we elect to not make use of chromosome maps. The summary of
  that additional text is that we do not belive we have an appropriate emough
  handle on the uncertanties in the chromosome maps to allow us to make
  statistically signifigant statements about population seperation. Due to the
  manner in which chromosome maps are constructed the fractional uncertanties
  of datapoints grow rapidly when compared to what they would be in on either a
  color-magnitude diagram or color-color diagram. We investigate this and found
  that the best-case (assuming no error on the red and blue fiducial line
  positioning and only possion error on each magnitude measurment) mean
  fractional error after computing a chromosome map was > 1. (See
  https://algebrist.com/~tboudreaux/files/ChromosomeMapFractionalErrorDist.pdf
  for the probability density distribution of fractional errors over the
  chromosome map parameter space). Because Fidanka fundamentally finds clusters
  based on a mean density derived from a monte carlo sampling of the error
  space, such large uncertanties prevent fidanka from identifying clusters
  reliably. CMD space does not face the same issues. It is true that we may be
  sacrificing some of the light-element abundance-variation driven seperation
  between potential stellar populations; however, as the aim of this work is to
  identify if chemically-self consistent modeling has a signifigant affect on
  the infered helium abundance it is not strictly nessisary to extract all
  previously identified populations. Finally, as there has been recent
  literature which questions whether there are in fact more than 2 populations
  in NGC 2808 (Valle et al. 2022, based on spectroscopy) we feel that it would
  be improper to not take an extremley careful handleing of clustering and
  the uncertanties which play into it.

Referee:
  Although the authors state several times that the work adopts a chemically
  self-consistent modeling of the globular cluster, it is not explained what
  the improvement is in this work compared to previous ones. Furthermore, it is
  not clear what the chemical composition of the two populations is (in terms
  of [Fe/H], [O/Fe], [N/Fe] etc.) and what the motivation behind this choice
  is. The authors refer to Milone et al. 2015 again but without further
  explanation and it is not up to the reader to go and read that paper. The
  paper must be comprehensive and all important information used must be
  clearly mentioned and reported. Finally, that paper provides the chemical
  content of the other three multiple populations (B, C and D). Why do the
  authors specifically use the chemical content of population E? The
  motivations for this are unclear, but very important.
Response:
  There has been text added to the manuscript to clarify what we mean by
  chemically self-consistent. To summarize here, we mean that unlike the
  majority of globular cluster modeling efforts we use high temperature
  opacities, low temperature opacities, atmosphereic bounday conditions, and
  structural compositons which all agree in their atmoic mass fractions up to
  Zinc (with heavier elements being excluded as the OPLIB high-temperature
  opacity tables we use have not been computed above zinc). This has been done,
  to the best of our knowledge, for only one other cluster, NGC 6752 (Dotter et
  al. 2015). That work showed that full chemical self-consistancy did not 
  result in a signifigant change to the infered helium abundances. Here we
  are essentially reproducing this work but for another cluster which has a more
  extreme infered helium abundance in the literature. We also include text in the
  paper to explain why we use populations A and in E in specific. These two populations
  have the lowest and highest infered helium abundances respectivley and have been 
  consitantly shown to be seperable in the NGC 2808 CMD. As we are focusing on 
  determining if chemically self-consistent modeling will affect the infered helium 
  abundance it makes sense to use the most extreme models. Finally, we have included
  [Fe/H], [O/Fe], and [N/Fe] values for the two populations which we do use in the
  text of the manuscript.

Referee:
  Similar to the previous comment, the authors do not mention the literature
  values adopted as first guess for the cluster age, distance and reddening.
  They do not comment on whether there is agreement between these values and
  the best-fit parameters. This, in my opinion, is fundamental to understand
  whether the best-fit solution obtained is reliable or not.
Response:
  Text has been added to the manuscript stating what the initial guesses were
  and how these values compare to the best fit parameters. The summary of this
  is that the best fit parameters are multiple sigma removed from the initial
  guess.

Referee:
  The best-fitting isochrones for Populations A and E in Table 1 have very
  different mixing length parameters (2.050 vs 1.600). This difference has a
  significant impact on the effective temperature (especially on the red giant
  branch) and is not physically justified in the text. In fact, there is no
  obvious reason to believe that two populations in the same cluster could have
  such different mixing lengths. However, this has a huge effect on the results
  obtained. Why didn't the authors set this parameter to be the same in both
  isochrones? What would have been the results in this case?
Response:
  TBD

Referee:
  Some of the Figures they present are not easily readable and understandable,
  of which Figure 6 is an example. It is not clear how to read the
  uncertainties on the different parameters from this figure (which are not
  even reported in Table 1). Furthermore, it is not explained in the text why
  the authors use the ACS artificial star tests to derive these uncertainties,
  when they could use the same Monte Carlo approach to make robust uncertainty
  predictions for each of the parameters.
Response:
  TBD

Referee:
  While reading the paper I noticed several typos, truncated sentences and few
  unanswered questions. The authors should read through the paper and clean up
  these issues.
Response:
  We have gone through the manuscript and further copy-edited it.