### 2024-03-22 ArXiv/conclusion.tex

\section{Conclusion}\label{sec:conclusion}
Here we have preformed the first chemically self-consistent modeling of the
Milky Way Globular Cluster NGC 2808. We find that, updated atmospheric boundary
conditions and opacity tables do not have a significant effect on the inferred
helium abundances of multiple populations. Specifically, we find that
population  has a helium mass fraction of 0.24, while population E has a helium
mass fraction of 0.39. Additionally, we find that the ages of these two populations 
agree within uncertainties. Further, we only find evidence for two distinct stellar
populations, which is in agreement with recent work studying the number
of populations in NGC 2808 spectroscopic data.

Further, we introduce a new software suite for globular cluster science,
\fidanka, which has been released under a permissive open source license.
\fidanka aims to provide a statistically robust set of tools for estimating the
parameters of multiple populations within globular clusters.